---
title: "MATH 216 Homework 4"
author: "TRISHA SINGH"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    smooth_scroll: false
---

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
# Suggested packages; feel free to use more!
library(tidyverse)
library(USAboundaries)
library(maptools)
library(sp)
library(broom)
library(readr)
library(rgeos)
library(rgdal)
library(ggplot2)
```



## Admistrative:

Please indicate

* Who you collaborated with: Connor, Yuchen
* Roughly how much time you spent on this HW so far: 2 hrs
* The URL of the RPubs published URL [here](http://rpubs.com/tsingh/hw4).
* What gave you the most trouble: Figuring out a set of good criteria for deciding the location of the shipping hub. There are too many things to take into consideration.
* Any comments you have:



## Question 1:

**Question**:  DHL, a German shipping company, has a ton of money to build a
shipping hub in the United States. Where should they build it?

**Deliverable**: A choropleth map of the lower 48 states (no Puerto Rico) where
the color corresponds to the distance of the given area to the shipping hub,
which is marked on the map with a large red dot.

```{r, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
# Getting US States:
counties_shp <- us_counties()

Census2010 <- read_csv("~/Documents/Fall 2016/Data Sc/HW-4/Census2010.csv") %>% 
  select_("Geo_FIPS", "SE_T001_001")

counties_data <- counties_shp@data %>% 
   mutate(centroid_long=coordinates(counties_shp)[,1]) %>% 
   mutate(centroid_lat=coordinates(counties_shp)[,2]) %>% 
  left_join(Census2010, by=c("geoid"="Geo_FIPS")) %>% 
  mutate(wt_centroid_long = SE_T001_001*centroid_long) %>% 
  mutate(wt_centroid_lat = SE_T001_001*centroid_lat)
 
counties_polygon <- tidy(counties_shp, region="geoid")

counties <- left_join(counties_polygon, counties_data, by=c("id"="geoid")) %>% 
   filter( !state_name %in% c("Alaska", "Hawaii", "Puerto Rico"))
  

#Calculate centre of population lat and long

centre_x_coord <- sum(counties_data$wt_centroid_long, na.rm=TRUE)/
  sum(counties_data$SE_T001_001, na.rm=TRUE)

centre_y_coord <- sum(counties_data$wt_centroid_lat, na.rm=TRUE)/
  sum(counties_data$SE_T001_001, na.rm=TRUE)

counties <- counties %>% 
  mutate(distance=sqrt((centre_x_coord-centroid_long)^2 +
                         (centre_y_coord-centroid_lat)^2))

cincinnati_x_coord <- -84.6630
cincinnati_y_coord <- 40.0533

chi_x_coord <--87.6298
chi_y_coord <- 41.8781

la_x_coord <- -118.2437
la_y_coord <- 34.0522

```


```{r, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
# Here is a template of a ggplot call; feel free to change this to suit your
# needs. Once your computations above are complete, set the eval=TRUE on the
# line above to tell R Markdown to run this code block.
ggplot(data=counties, aes(x=long, y=lat, group=group, fill=distance)) +
  # Plot choropleth polygons, low -> high being colored white -> dark green
  geom_polygon() +
  scale_fill_gradient(low="white", high="darkgreen") + 
  # Trace outlines of areas
  geom_path(size=0.1) +
  # Mark hub
  annotate("text", label="Centre of Population", 
           x=centre_x_coord, y=centre_y_coord+0.5, col="red", size=5) +
  annotate("point", 
           x=centre_x_coord, y=centre_y_coord, col="red", size=2) +
  annotate("text", label="Existing DHL Cincinnati Airport Hub", 
           x=cincinnati_x_coord, y=cincinnati_y_coord+0.5, col="red", size=5) +
  annotate("point", 
           x=cincinnati_x_coord, y=cincinnati_y_coord, col="red", size=2) +
  annotate("text", label="Potential Chicago Hub", 
           x=chi_x_coord, y=chi_y_coord+0.5, col="blue", size=5) +
  annotate("point", 
           x=chi_x_coord, y=chi_y_coord, col="blue", size=2) +
  annotate("text", label="Potential LA Hub", 
           x=la_x_coord, y=la_y_coord+0.5, col="blue", size=5) +
  annotate("point", 
           x=la_x_coord, y=la_y_coord, col="blue", size=2) +
  coord_map()
```

The goal is to locate the new shipping hub as close to the centre of population as possible, taking other factors into consideration.

* DHL already has a shipping hub at Cincinnati Airport which it is planning to expand, so rather than international flight shipping, we should look at hubs with good ground and/or water transport.
Citation: [here](http://www.wsj.com/articles/dhl-to-spend-108-million-to-expand-cincinnati-air-hub-1433101099)

* The top locations for shipping hubs with the best ground and water transport are: _Chicago_, _LA_, Houston, Memphis, Atlanta
Citation: [here](http://www.supplychaindigital.com/logistics/2913/Top-5-US-logistics-cities)

* I am not going to consider Houston because it is a booming trading hub due to the North American Free Trade Association (NAFTA), which may get scrapped under the Trump government. I will also not consider Memphis since FedEx's shipping hub is located there and it could not be a good idea to compete with FedEx for supply chain resources. I am exclusing Atlanta because it does not have a water port.

**So Chicago and LA are good options for opening a shipping hub. Chicago is close to the centre of population, but it is also close to the existing hub. LA is far from the centre of population, but it covers a side of the US that is not already covered by DHL.**




